---
title: "Week 5 R Fundamentals"
subtitle: "Lab 3 Caloric Intake and Demographics"
author: "Tony Lim"
output:
  html_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Overview

The goal of this analysis is to explore the relationship between total caloric intake, including beer consumption, and demographic factors. We will utilize some important R tools, including data importing, merging, summarizing by groups, and linear and log-linear regression.

# 2. Data Collection and Merging
 
We will use 2 datasets from the National Health and Nutrition Examination Survey (NHANES) for 2015-2016. Note, the dataset TOTAL_CALORIES was previously created using the NHANES datasets DR1IFF_I.XPT (on daily food consumption) and DRXFCD_I.XPT (on food codes) available at https://wwwn.cdc.gov/nchs/nhanes/Default.aspx.

**DEMO** 

* **SEQN:**          Survey respondant's sequence number
* **AGE:**           Age of respondant (years)
* **GENDER:**        Biological sex of respondant
* **EDUCATION:**     Education level of respondant ("Less than High School", "High School", "Some College", "College Grad",)

**TOTAL_CALORIES**

* **SEQN:**          Survey respondant's sequence number
* **ALL_CALORIES:**  Total calories consumed (food + drinks) by respondant in past 24 hours     
* **BEER_CALORIES:** Calories consumed from beer in past 24 hours

Use `read.csv()` to load each dataset for analysis.
```{r}
library(readr)

DEMO <- read.csv("DEMO.csv")

TOTAL_CALORIES <- read.csv("TOTAL_CALORIES.csv")
```

## Question 1

After loading the dplyr library, use `full_join()` to merge the datasets using `SEQN` as the common link. Name this new dataset FULL. How many observations are in FULL?

```{r}
library(dplyr)

FULL  <- full_join(DEMO, TOTAL_CALORIES, by = "SEQN")
```

**SOLUTION:** With `full_join()`, we have 9,207 observations.

## Question 2

Use `inner_join()` to merge the datasets using `SEQN` as the common link. Name this new dataset INNER. How many observations are in INNER?

```{r}
INNER <- inner_join(DEMO, TOTAL_CALORIES, by = "SEQN")
```

**SOLUTION:** With `inner_join()`, we have 5,017 observations.

We will use INNER in the remainder of the lab, as we want only to include only those individuals with both demographic information and calorie consumption data.

# 3. Data Exploration

Let's calculate the mean and standard deviation of TOTAL calories by gender using the following code:

```{r}
tapply(INNER$ALL_CALORIES, INNER$GENDER, mean)

tapply(INNER$ALL_CALORIES, INNER$GENDER, sd)

```

Women have an average total caloric consumption of 1799.3 with a standard deviation of 761.1. Men have an average total caloric consumption of 2373.9 with a standard deviation of 1039.3. 

## Question 3

What are the mean BEER calories consumed by men and women?
```{r}
tapply(INNER$BEER, INNER$GENDER, mean)
```
**SOLUTION:** On average, men consume 92.4 beer calories while women consume 16.6 beer calories.  

## Question 4

Consumers with which education level consume the most TOTAL calories? And the most BEER calories?
```{r}
tapply(INNER$ALL_CALORIES, INNER$EDUCATION, mean)
```

```{R}
tapply(INNER$BEER_CALORIES, INNER$EDUCATION, mean)
```

**SOLUTION:** On average, college graduates consume the most total calories (2119.7), but individuals with a high school degree consume the most beer calories (77.7).

# 4. Data Visualization

## Question 5

Create a histogram of total caloric consumption by gender, using `ggplot()` and `facet_wrap()` with different colors.

```{r}
library(ggplot2)

INNER %>%
  ggplot() +
  geom_histogram(mapping = aes(x = ALL_CALORIES, fill = GENDER), color = "white") +
  labs(title = "Distribution of total caloric consumption by gender",
       x = "Total caloric consumption",
       y = "Count") +
  scale_fill_manual(name = "Gender", values = c("#2774AE", "#FFD100")) +
  facet_wrap(~ GENDER)
```

# 5. Linear Regression Analysis

## Question 6 

Regress TOTAL calories on age, gender, and education using the INNER data. 

```{r}
library(jtools)

reg1 <- lm(ALL_CALORIES ~ AGE + GENDER+ EDUCATION, data = INNER)

summ(reg1)
```

## Question 7

Regress BEER calories on age, gender, and education using the INNER data.

```{r}
reg2 <- lm(BEER_CALORIES ~ AGE + GENDER + EDUCATION, data = INNER)

summ(reg2)
```

Finally, let's use our regressions to make a prediction for a friend. We make a small dataframe called FRIEND and enter some input values for the regression.

```{r}
FRIEND <- data.frame(
  AGE = 25,
  GENDER = "Male",
  EDUCATION = "College Grad"
)
```

Then we enter this information into the `predict()` command. We can also obtain a 95% prediction interval.

```{r}
predict(reg1, newdata = FRIEND, interval = "prediction")
```

## Question 8

Your neighbor is an Instagram influencer who is 32, attended some college, and is female. What is her predicted calorie consumption?

```{r}
NEIGHBOR <- data.frame(
  AGE = 32,
  GENDER = "Female",
  EDUCATION = "Some College"
)

predict(reg1, newdata = NEIGHBOR, interval = "prediction")
```
