---
title: "Final Project Predicting Coffee Consumption Using Multiple Linear Regression"
author: "Team 6 (Alex Frumkin, Tony Lim, Sara Nager, Jaanhvi Vaidya, Trisha Mathelier)"
output:
  html_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

For our final project, we wanted to analyze coffee consumption (measured in grams) and to determine whether a multiple linear regression was applicable. We utilized datasets from the [National Health and Nutrition Examination Survey (NHANES) 2015-2016](https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2015). 

Specifically, we used

* [Demographics Data](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DEMO_I.htm) ([DEMO_I.XPT](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Demographics&CycleBeginYear=2015))
* [Dietary Interview - Individual Foods, First Day](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR1IFF_I.htm) ([DR1IFF_I.XPT](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Dietary&CycleBeginYear=2015))
* [Dietary Interview Technical Support File - Food Codes](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DRXFCD_I.htm) ([DRXFCD_I.XPT](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Dietary&CycleBeginYear=2015))

Variables in the `final` dataset include:

* **SEQN**: Respondent sequence number
* **TOTAL**: Total number of calories consumed
* **COFFEE_TOTAL**: Total number of grams attributable to coffee consumption
* **GENDER**: Gender
* **AGE**: Age in years at screening
* **RACE**: Race
    + Black
    + Hispanic
    + White
    + Other/Multi
* **EDUCATION**: Education level
    + Less than high school
    + Some high school
    + High school graduate/GED
    + Some college/AA
    + College graduate or above
    + Unsure
* **MARITAL**: Marital status
    + Married
    + Widowed
    + Divorced
    + Separated
    + Never married
    + Living with partner
    + Refused
    + Unsure
* **HH_SIZE**: Total number of people in the household
* **HH_INC**: Total household income

# 1. Installing Libraries

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(foreign)
library(ggplot2)
library(jtools)
library(ggpubr)
library(tidyverse)
library(wesanderson)
```

# 2. Data Preparation
## Demographic Dataset

```{r}
demo <- read.xport("DEMO_I.XPT")

(demo <- as_tibble(demo))
```

Converting DMDEDUC2 variable for those below 20 years of age to DMDEDUC3 variable
```{r}
demo$DMDEDUC2[demo$DMDEDUC3 < 10] <- 1
demo$DMDEDUC2[demo$DMDEDUC3 %in% c(9:12)] <- 2
demo$DMDEDUC2[demo$DMDEDUC3 %in% c(13, 14)] <- 3
demo$DMDEDUC2[demo$DMDEDUC3 == 15] <- 4
demo$DMDEDUC2[demo$DMDEDUC3 %in% c(55, 66)] <- 1
demo$DMDEDUC2[demo$DMDEDUC3 == 77] <- 7
demo$DMDEDUC2[demo$DMDEDUC3 == 99] <- 9
```

```{r, echo = FALSE, include = FALSE}
# Filtering out people younger than 20 years old because there are two education variables (one for those up to 19 years of age and one for those 20 years of age and older).
# (demo <- demo %>%
#   filter(RIDAGEYR >= 20))
```

Selecting the desired variables and renaming them. 
```{r}
(demo <- demo %>%
  select(SEQN, GENDER = RIAGENDR, AGE = RIDAGEYR, RACE = RIDRETH1, EDUCATION = DMDEDUC2, MARITAL = DMDMARTL, HH_SIZE = DMDHHSIZ,
         HH_INC = INDHHIN2))
```

## Factorizing Variables

```{r}
demo$AGE <- cut(demo$AGE, 
                 breaks = c(0, 15, 25, 35, 45, 55, 65, 75, Inf), 
                 labels = c("Below 15", "15-25", "25-35", "35-45", "45-55", "55-65", "65-75", "Above 75"))

demo$GENDER <- factor(demo$GENDER, 
                       labels = c("Male", "Female"))

demo$EDUCATION <- factor(demo$EDUCATION, 
                        labels = c("Less than high school", "Some high school", "High school graduate/GED", "Some college/AA", "College graduate or above", "Unsure"))

demo$RACE <- factor(demo$RACE, 
                     labels=c("Hispanic", "Hispanic", "White", "Black", "Other/Multi"))

# Filtered out 12 ($20,000 and Over) and 13 (Under $20,000)
demo <- demo %>%
  filter(!(HH_INC == 12 | HH_INC == 13))

demo$MARITAL <- factor(demo$MARITAL, 
                        labels = c("Married", "Widowed", "Divorced", "Separated", "Never married", "Living with partner", "Refused", "Unsure"))

demo$HH_INC <- cut(demo$HH_INC,
                    breaks = c(0, 4.5, 7.5, 10.5, 76, 98, Inf), 
                    labels = c("Below 20k", "20-45k", "45-75k", "Above 75k", "Refused", "Unsure"))

demo %>%
  print(width = Inf)
```

## Joining the Dietary and Food Code Datasets - `join`

Reading in the dietary dataset and selecting/renaming the desired variables.
```{r}
diet <- read.xport("DR1IFF_I.XPT")

(diet <- as_tibble(diet) %>%
  select(SEQN, KCAL = DR1IKCAL, GRAM = DR1IGRMS, USDA = DR1IFDCD))
```

Reading in the food code dataset and renaming the variables.
```{r}
food_code <- read.xport("DRXFCD_I.XPT")

(food_code <- as_tibble(food_code) %>%
  select(USDA = DRXFDCD, SHORT = DRXFCSD, LONG = DRXFCLD))
```

Combining the dietary dataset with the food code dataset to create `join`.
```{r}
join <- inner_join(diet, food_code, by = "USDA")

(join <- as_tibble(join))
```

## Calculating Total Calories - `total`

Calculating the total number of calories consumed.
```{r}
total <- join %>%
  group_by(SEQN) %>%
  mutate(TOTAL = sum(KCAL)) %>%
  select(SEQN, TOTAL)

(total <- total %>% 
   distinct(SEQN, TOTAL))
```

## Calculating Total Grams of Coffee - `coffee`

Selecting the relevant food codes such as coffee creamer and multiple forms of coffee (brewed, cappuccino, Cuban, espresso, iced, instant, latte, macchiato, etc.) while ommitting unrelated coffee products (coffee cake, Irish coffee, etc.).
```{r}
coffee_code <- food_code[c(393:405, 8254:8374, 8378), ]
(coffee_code <- as_tibble(coffee_code))
```

Finding the total number of grams attributable to consuming coffee or coffee-related products.
```{r}
coffee <- inner_join(join, coffee_code, by = "USDA")

coffee <- coffee %>%
  group_by(SEQN) %>%
  mutate(COFFEE_TOTAL = sum(GRAM)) %>%
  select(SEQN, COFFEE_TOTAL)

(coffee <- coffee %>%
  distinct(SEQN, COFFEE_TOTAL))
```

## Final Dataset - `final`

First, we created `clean`, a dataset with `SEQN`, `TOTAL`, `COFFEE_TOTAL` using a left join between `total` and `coffee`.
```{r}
(clean <- left_join(total, coffee, by = "SEQN"))
```

Then we created `final` using inner join `clean` with `demo`. 
```{r}
final <- inner_join(clean, demo, by = "SEQN")

final$COFFEE_TOTAL <- replace_na(final$COFFEE_TOTAL, 0)

(final <- as_tibble(final))
```

# 3. Data Visualization

## Age distribution
```{r, message = FALSE, warning = FALSE}
age.p <- final %>%
  filter(!is.na(AGE)) %>%
  ggplot(mapping = aes(x = AGE)) + 
  geom_bar(fill = "#fff1d8", color = "white") +
  labs(title = "Age distribution",
       x = "Age",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Education distribution
```{r, message = FALSE, warning = FALSE}
education.p <- final %>%
  ggplot(mapping = aes(x = EDUCATION)) + 
  geom_bar(fill = "#a06d4c", color = "white") +
  labs(title = "Education distribution",
       x = "Education",
       y = "Count") +
  theme_minimal() +
  coord_flip() + 
  scale_x_discrete(limits = rev(levels(final$EDUCATION)))
```

## Race distribution
```{r, message = FALSE, warning = FALSE}
race.p <- final %>%
  ggplot(mapping = aes(x = RACE)) + 
  geom_bar(fill = "#dc9265", color = "white") +
  labs(title = "Race distribution",
       x = "Race",
       y = "Count") +
  theme_minimal()
```

## Household income distribution
```{r, message = FALSE, warning = FALSE}
hh_inc.p <- final %>%
  ggplot(mapping = aes(x = HH_INC)) + 
  geom_bar(fill = "#ee7936", color = "white") +
  labs(title = "Household income distribution",
       x = "Household income",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r, message = FALSE, warning = FALSE}
ggarrange(age.p,  hh_inc.p,
          race.p, education.p,
          ncol = 2, 
          nrow = 2)
```

## Household size distribution
```{r, message = FALSE, warning = FALSE}
final %>%
  ggplot(mapping = aes(x = HH_SIZE)) + 
  geom_bar(fill = "#D69C4E", color = "white") +
  labs(title = "Household size distribution",
       x = "Household size",
       y = "Count") +
  scale_x_discrete(limits = c(1:7)) +
  theme_minimal()
```

```{r, message = FALSE, warning = FALSE}
total.p <- final %>%
  ggplot(mapping = aes(x = TOTAL)) + 
  geom_histogram(fill = "#a06d4c", color = "white") +
  labs(title = "Distribution of calories",
       x = "Calories",
       y = "Count") +
  theme_minimal()
```

```{r}
total_gender.p <- final %>%
  ggplot(mapping = aes(x = TOTAL, fill = GENDER)) + 
  geom_histogram(color = "white") +
  labs(title = "Distribution of calories by gender",
       x = "Calories",
       y = "Count") +
  theme_minimal() +
  scale_fill_manual(name = "Gender", values = c("#ee7936", "#599ed6")) +
  facet_wrap(~ GENDER)
```

```{r, message = FALSE, warning = FALSE}
ggarrange(total.p,
          total_gender.p,
          ncol = 1)
```


## Distribution of coffee consumption
```{r, message = FALSE, warning = FALSE}
coffee_total.p <- final %>%
  ggplot(mapping = aes(x = COFFEE_TOTAL)) + 
  geom_histogram(fill = "#a06d4c", color = "white") +
  labs(title = "Distribution of coffee consumption",
       x = "Coffee consumption (grams)",
       y = "Count") +
  theme_minimal()
```

## Distribution of coffee consumption by gender
```{r, message = FALSE, warning = FALSE}
coffee_total_gender.p <- final %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = COFFEE_TOTAL, fill = GENDER), color = "white") +
  labs(title = "Distribution of coffee consumption by gender",
       x = "Coffee consumption (grams)",
       y = "Count") +
  scale_fill_manual(name = "Gender", values = c("#ee7936", "#599ed6")) +
  facet_wrap(~ GENDER) +
  theme_minimal()
```

```{r, message = FALSE, warning = FALSE}
ggarrange(coffee_total.p,
          coffee_total_gender.p,
          ncol = 1)
```

```{r}
avgcoffee_education.p <- final %>%
  filter(!is.na(EDUCATION)) %>%
  group_by(EDUCATION) %>%
  mutate(COFFEE_EDUCATION = mean(COFFEE_TOTAL)) %>%
  select(EDUCATION, COFFEE_EDUCATION) %>%
  distinct(EDUCATION, COFFEE_EDUCATION) %>%
  ggplot() + 
  geom_bar(mapping = aes(x = EDUCATION, y = COFFEE_EDUCATION), 
           fill = "#a06d4c", 
           stat = "identity") +
  labs(title = "Average coffee consumption by education",
       x = "Education",
       y = "Average coffee consumption (grams)") +
  theme_minimal() +
  coord_flip() + 
  scale_x_discrete(limits = rev(levels(final$EDUCATION)))
```

```{r}
avgcoffee_race.p <- final %>%
  #filter(!is.na(EDUCATION)) %>%
  group_by(RACE) %>%
  mutate(COFFEE_RACE = mean(COFFEE_TOTAL)) %>%
  select(RACE, COFFEE_RACE) %>%
  distinct(RACE, COFFEE_RACE) %>%
  ggplot() + 
  geom_bar(mapping = aes(x = RACE, y = COFFEE_RACE), 
           fill = "#dc9265", 
           stat = "identity") +
  labs(title = "Average coffee consumption by race",
       x = "Race",
       y = "Average coffee consumption (grams)") +
  theme_minimal()
```

```{r, message = FALSE, warning = FALSE}
ggarrange(avgcoffee_education.p,
          avgcoffee_race.p,
          ncol = 1)
```

## Distribution coffee consumption by age
```{r, message = FALSE, warning = FALSE}
final %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = COFFEE_TOTAL), color = "white") +
  labs(title = "Distribution of coffee consumption by age",
       x = "Coffee consumption (grams)",
       y = "Count") +
  #scale_fill_manual(name = "Coffee consumption (grams)", values = wes_palette("Moonrise2")) +
  facet_wrap(~ AGE) +
  theme_minimal()
```

## Distribution of coffee consumption by race
```{r, message = FALSE, warning = FALSE}
final %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = COFFEE_TOTAL, fill = RACE), color = "white") +
  labs(title = "Distribution of coffee consumption by race",
       x = "Coffee consumption (grams)",
       y = "Count") +
  scale_fill_manual(name = "Race", values = wes_palette("Moonrise2")) +
  facet_wrap(~ RACE) +
  theme_minimal()
```

## Distribution of coffee consumption by education
```{r, message = FALSE, warning = FALSE}
final %>%
  filter(!(EDUCATION == "Unsure" | is.na(EDUCATION))) %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = COFFEE_TOTAL, fill = EDUCATION), color = "white") +
  labs(title = "Distribution of coffee consumption by education",
       x = "Coffee consumption (grams)",
       y = "Count") +
  scale_fill_manual(name = "Education", values = wes_palette("IsleofDogs2")) +
  facet_wrap(~ EDUCATION) +
  theme_minimal()
```

# 4. Data Analysis

```{r}
summary(final)
```

```{r}
final[which.max(final$COFFEE_TOTAL), ]
```

```{r}
tapply(final$TOTAL, final$GENDER, mean, na.rm = T)
```

```{r}
tapply(final$COFFEE_TOTAL, final$GENDER, mean, na.rm = T)
tapply(final$COFFEE_TOTAL, final$GENDER, median)
```

```{r}
tapply(final$TOTAL, final$EDUCATION, mean, na.rm = T)
```

```{r}
tapply(final$COFFEE_TOTAL, final$EDUCATION, mean, na.rm = T)
```

```{r}
tapply(final$TOTAL, final$RACE, mean, na.rm = T)
```

```{r}
tapply(final$COFFEE_TOTAL, final$RACE, mean, na.rm = T)
```

```{r}
final %>%
  t.test(COFFEE_TOTAL ~ GENDER, data = ., var.equal = TRUE)
```

# 5. Multiple Linear Regression

```{r}
reg <- glm(COFFEE_TOTAL ~ GENDER + AGE + RACE + EDUCATION + MARITAL + HH_SIZE + HH_INC, data = final)
summary(reg)
```

```{r}
reg2 <- glm(COFFEE_TOTAL ~ GENDER + AGE + RACE + EDUCATION + HH_SIZE, data = final)
summary(reg2)
```

```{r}
tony <- data.frame(
  GENDER = "Male",
  AGE = "15-25",
  RACE = "Other/Multi",
  EDUCATION = "College graduate or above",
  HH_SIZE = 1
  )
```

```{r}
predict(reg2, newdata = tony)
```

# Bonus

Finding the coffee products with the highest average calories. 
```{r}
diet <- read.xport("DR1IFF_I.XPT")

diet <- diet %>%
  select(SEQN, KCAL = DR1IKCAL, GRAMS = DR1IGRMS, USDA = DR1IFDCD)

avgkcal <- right_join(diet, coffee_code, by = "USDA") %>%
  group_by(USDA) %>%
  mutate(AVGKCAL = mean(KCAL / GRAMS) * mean(GRAMS)) %>%
  select(USDA, AVGKCAL, SHORT, LONG) %>%
  distinct(.) %>%
  arrange(-(AVGKCAL))

right_join(diet, coffee_code, by = "USDA") %>%
  group_by(USDA) %>%
  mutate(AVGKCAL = mean(KCAL)) %>%
  select(USDA, AVGKCAL, SHORT, LONG) %>%
  distinct(.) %>%
  arrange(-(AVGKCAL))
```