---
title: "Final Project"
author: "Team 6 (Alex Frumkin, Tony Lim, Sara Nager, Jaanhvi Vaidya, Trisha Mathelier)"
output:
  html_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For our final project, we wanted to analyze the calories attributed to coffee consumption and determine whether a linear regression is applicable. We used datasets from the [National Health and Nutrition Examination Survey (NHANES) 2015-2016](https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2015). 

Specifically, we used

* [Demographics Data](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DEMO_I.htm) ([DEMO_I.XPT](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Demographics&CycleBeginYear=2015))
* [Dietary Interview - Individual Foods, First Day](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR1IFF_I.htm) ([DR1IFF_I.XPT](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Dietary&CycleBeginYear=2015))
* [Dietary Interview Technical Support File - Food Codes](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DRXFCD_I.htm) ([DRXFCD_I.XPT](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Dietary&CycleBeginYear=2015))

Variables in the `final` dataset include:

* **SEQN**: Respondent sequence number
* **TOTAL**:
* **COFFEE_TOTAL**: 
* **GENDER**: Gender
* **AGE**: Age in years at screening
* **RACE**: Race
    + Black
    + Hispanic
    + White
    + Other/Multi
* **EDUCATION**: Education level for adults (20+ years of age)
    + Less than high school
    + Some high school
    + High school graduate/GED
    + Some college/AA
    + College graduate or above
    + Unsure
* **MARITAL**: Marital status
    + Married
    + Widowed
    + Divorced
    + Separated
    + Never married
    + Living with partner
    + Refused
    + Unsure
* **HH_SIZE**: Total number of people in the household
* **HH_INC**: Total household income

# 1. Installing Libraries

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(foreign)
library(ggplot2)
library(jtools)
library(tidyverse)
library(wesanderson)
```

# 2. Data Preparation
## Demographic Dataset

```{r}
DEMO_I <- read.xport("DEMO_I.XPT")

(DEMO_I <- as_tibble(DEMO_I))
```

```{r}
DEMO_I$DMDEDUC2[DEMO_I$DMDEDUC3 < 10] <- 1
DEMO_I$DMDEDUC2[DEMO_I$DMDEDUC3 %in% c(9:12)] <- 2
DEMO_I$DMDEDUC2[DEMO_I$DMDEDUC3 %in% c(13, 14)] <- 3
DEMO_I$DMDEDUC2[DEMO_I$DMDEDUC3 == 15] <- 4
DEMO_I$DMDEDUC2[DEMO_I$DMDEDUC3 %in% c(55, 66)] <- 1
DEMO_I$DMDEDUC2[DEMO_I$DMDEDUC3 == 77] <- 7
DEMO_I$DMDEDUC2[DEMO_I$DMDEDUC3 == 99] <- 9
```

Filtering out people younger than 20 years old because there are two education variables (one for those up to 19 years of age and one for those 20 years of age and older).

```{r}
# (demo <- DEMO_I %>%
#   filter(RIDAGEYR >= 20))
```

Selecting the desired variables and renaming them. 

```{r}
(demo <- DEMO_I %>%
  select(SEQN, GENDER = RIAGENDR, AGE = RIDAGEYR, RACE = RIDRETH1, EDUCATION = DMDEDUC2, MARITAL = DMDMARTL, HH_SIZE = DMDHHSIZ,
         HH_INC = INDHHIN2))
```

## Factorizing Variables

```{r}
demo$AGE <- cut(demo$AGE, 
                 breaks = c(0, 25, 35, 45, 55, 65, 75, Inf), 
                 labels = c("Below 25", "25-35", "35-45", "45-55", "55-65", "65-75", "Above 75"))

demo$GENDER <- factor(demo$GENDER, 
                       labels = c("Male", "Female"))

demo$EDUCATION <- factor(demo$EDUCATION, 
                        labels = c("Less than high school", "Some high school", "High school graduate/GED", "Some college/AA", "College graduate or above", "Unsure"))

demo$RACE <- factor(demo$RACE, 
                     labels=c("Hispanic", "Hispanic", "White", "Black", "Other/Multi"))

# Filtered out 12 ($20,000 and Over) and 13 (Under $20,000)
demo <- demo %>%
  filter(!(HH_INC == 12 | HH_INC == 13))

demo$MARITAL <- factor(demo$MARITAL, 
                        labels = c("Married", "Widowed", "Divorced", "Separated", "Never married", "Living with partner", "Refused", "Unsure"))

demo$HH_INC <- cut(demo$HH_INC,
                    breaks = c(0, 4.5, 7.5, 10.5, 76, 98, Inf), 
                    labels = c("Below 20k", "20-45k", "45-75k", "Above 75k", "Refused", "Unsure"))

demo %>%
  print(width = Inf)
```

## Combined Dataset

Combining the first day dataset with the food code dataset to create `join`.

```{r}
DR1IFF_I <- read.xport("DR1IFF_I.XPT")

# Rename column to the same column in DRXFCD_I to inner_join later
DR1IFF_I <- DR1IFF_I %>%
  rename(DRXFDCD = DR1IFDCD)

DRXFCD_I <- read.xport("DRXFCD_I.XPT")
```

```{r}
join <- inner_join(DR1IFF_I, DRXFCD_I, by = "DRXFDCD")
```

## Total Calories Dataset

Finding the total number of calories consumed for each sample. 

```{r}
total <- DR1IFF_I %>%
  group_by(SEQN) %>%
  mutate(TOTAL = sum(DR1IKCAL)) %>%
  select(SEQN, TOTAL)

(total <- total %>% 
   distinct(SEQN, TOTAL))
```

## Total Coffee Calories Dataset

Selecting the relevant food codes such as coffee creamer and multiple forms of coffee (brewed, cappuccino, Cuban, espresso, iced, instant, latte, macchiato, etc.) while ommitting unrelated coffee products (coffee cake, Irish coffee, etc.).

```{r}
coffee_code <- DRXFCD_I[c(393:405, 8254:8374, 8378), ]
(coffee_code <- as_tibble(coffee_code))
```

Finding the total number of calories attributed to consuming coffee or coffee-related products.

```{r}
coffee <- inner_join(join, coffee_code, by = "DRXFDCD")

coffee <- coffee %>%
  group_by(SEQN) %>%
  mutate(COFFEE_TOTAL = sum(DR1IKCAL)) %>%
  select(SEQN, COFFEE_TOTAL)

(coffee <- coffee %>%
  distinct(SEQN, COFFEE_TOTAL))
```

## Final Dataset

First, we created `clean`, a dataset with `SEQN`, `TOTAL`, `COFFEE_TOTAL` using a left join between `total` and `coffee`. Then we created `final` using inner join `clean` with `demo`. 

```{r}
(clean <- left_join(total, coffee, by = "SEQN"))

final <- inner_join(clean, demo, by = "SEQN")

final$COFFEE_TOTAL <- replace_na(final$COFFEE_TOTAL, 0)

(final <- as_tibble(final))
```

# 3. Data Visualization

```{r, message = FALSE, warning = FALSE}
final %>%
  ggplot(mapping = aes(x = COFFEE_TOTAL)) + 
  geom_histogram(fill = "#D69C4E", color = "white") +
  labs(title = "Distribution of calories from coffee consumption",
       x = "Coffee consumption (calories)",
       y = "Count")
```

```{r, message = FALSE, warning = FALSE}
final %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = COFFEE_TOTAL, fill = GENDER), color = "white") +
  labs(title = "Distribution of calories from coffee consumption",
       x = "Coffee consumption (calories)",
       y = "Count") +
  scale_fill_manual(name = "Gender", values = wes_palette("Darjeeling2")) +
  facet_wrap(~ GENDER)
```

```{r, message = FALSE, warning = FALSE}
final %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = COFFEE_TOTAL, fill = RACE), color = "white") +
  labs(title = "Distribution of calories from coffee consumption",
       x = "Coffee consumption (calories)",
       y = "Count") +
  scale_fill_manual(name = "Race", values = wes_palette("Moonrise2")) +
  facet_wrap(~ RACE)
```

```{r, message = FALSE, warning = FALSE}
final %>%
  filter(!(EDUCATION == "Unsure" | is.na(EDUCATION))) %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = COFFEE_TOTAL, fill = EDUCATION), color = "white") +
  labs(title = "Distribution of calories from coffee consumption",
       x = "Coffee consumption (calories)",
       y = "Count") +
  scale_fill_manual(name = "Education", values = wes_palette("IsleofDogs2")) +
  facet_wrap(~ EDUCATION)
```

# 4. Data Analysis

```{r}
summary(final)
```


```{r}
final[which.max(final$COFFEE_TOTAL), ]
```

```{r}
tapply(final$TOTAL, final$GENDER, mean, na.rm = T)
```

```{r}
tapply(final$COFFEE_TOTAL, final$GENDER, mean, na.rm = T)
```

```{r}
tapply(final$TOTAL, final$EDUCATION, mean, na.rm = T)
```

```{r}
tapply(final$COFFEE_TOTAL, final$EDUCATION, mean, na.rm = T)
```

```{r}
tapply(final$TOTAL, final$RACE, mean, na.rm = T)
```

```{r}
tapply(final$COFFEE_TOTAL, final$RACE, mean, na.rm = T)
```

```{r}
final %>%
  t.test(COFFEE_TOTAL ~ GENDER, data = ., var.equal = TRUE)
```

# 5. Linear Regression

```{r}
reg <- glm(COFFEE_TOTAL ~ GENDER + AGE + RACE + EDUCATION + MARITAL + HH_SIZE + HH_INC, data = final)
summary(reg)
```

```{r}
reg2 <- glm(COFFEE_TOTAL ~ GENDER + AGE + RACE + EDUCATION, data = final)
summary(reg2)
```

```{r}
tony <- data.frame(
  GENDER = "Male",
  AGE = "Below 25",
  RACE = "Other/Multi",
  EDUCATION = "College graduate or above"
  )
```

```{r}
predict(reg2, newdata = tony, interval = "prediction")
```


